### 2013-09-19 intro.tex
say everything twice. comit and delete lol

%\includegraphics[width=\linewidth]{python_soul}%
%\flushright%
%Randall Munroe, \emph{xkcd.com}
\end{minipage}%
\\[2cm]
In computer vision research, the goal is to automatically extract information from an image.
In particular discerning semantic information, that is interpreting an image, is a prominent
research topic.
While much process has been made in recent years, computer vision systems still lack behind
humans in most tasks that require semantic information. Many semantic tasks can be formulated
in terms of \emph{semantic classes}, that is categories of parts, objects or scenes.
Examples include answering questions such as ``Is this a picture of a beach?'', ``How many
cars are in this image'' or even ``What objects lie on the table?''.
These questions illustrate a range of possible tasks involving semantic categories,
from classifying images of single objects, over localization and counting of object classes,
to parsing a scene fully into objects and object classes, together with their relations.
%
While humans can distinguish tens of thousands of object classes, and have little trouble
in interpreting complex scenes, current methods are often restricted to a much smaller number
of classes. While some recent approaches are able to classify images into a thousand 
or even ten thousand classes, results are still somewhat mixed.
In particular, when refining the task to localization, and allowing multiple objects within
a scene, few approaches have been evaluated on more than about twenty classes.

In this work, we will investigate the tasks of semantic segmentation and object
class segmentation.  These ask for a semantic annotation of each pixel with one
of a set of known classes.

We choose this task in particular for several reasons:
\begin{itemize}
    \item Pixel-level annotation provides highly detailed information about the scene.
    \item Joint estimation of multiple classes allows for the use of \emph{context}.
    \item In contrast to category-agnostic segmentation approaches, object class segmentation
        has an unambiguous true labeling.
    \item A variety of annotated datasets is publicly available.
\end{itemize}
%
Applications of semantic segmentation and scene understanding include automatic
image interpretation for retrieval, autonomous driving and mobile robotics.
While these are already important applications, we expect that with the
abundance of camera data, and the proliferation of mobile computing, even
more applications will gain importance soon.

We will in the following distinguish between the task of semantic segmentation,
which is usually distinguishing unstructured ``stuff'' classes such as road and
grass, and object class segmentation, which denotes the segmentation of very
structured classes, such as cars, planes and people.
We consider four different datasets in this thesis: the object class segmentation
datasets Graz02 and Pascal VOC 2010, and the semantic segmentation datasets MSRC-21,
and NYU V2.
%
Both tasks have the same ultimate goal of parsing, and therefore understanding, images
in terms of semantic classes, but usually employ  different mechanisms to represent the input.

One of the bottlenecks in learning object class segmentation and semantic
segmentation is the availability of training data.  While unlabeled image data,
and even data with semantic
``tags'' is available in practically unlimited quantities, semantic annotation
on pixel level is only available through laborious manual annotation. 
Chapter %FIXME
introduces an approach to cope with this shortcoming of object class segmentation approaches
by introducing a method to find segmentations automatically from image level annotations.

@@

\citet{nowozin2010parameter} provide a detailed evaluation of different aspects of learning object class
segmentation, that is somewhat orthogonal to this work. Their work considers the choice of features,
number superpixels and pairwise potentials for conditional maximum likelihood learning of tree-structured CRFs.
\citet{nowozin2010parameter} also compare conditional maximum likelihood learning with maximum margin learning,
finding little difference in accuracy.
We focus our work on the more popular neighborhood graphs, which do not allow for easy inference.
Therefore, conditional maximum likelihood learning would be much harder in our setting. For our studies,
we fix the considered feature set, and vary the maximum margin learning algorithm. Despite
this seemingly limited setup, we can make out large differences in the applicability of
some of the commonly used methods. For comparison, we also evaluate the learning algorithms
on chain CRFs, where inference is simple.

%related work?? here again?
%Current systems for semantic segmentation emply a wealth of different paradigms, most notably:
%The ranking of figure ground hypotheses
%conditional random field / superpixel? kraehenbuehl?
%detector based
%sparse coding

%
%Structured approaches
%

>>>>

%\includegraphics[width=\linewidth]{python_soul}%
%\flushright%
%Randall Munroe, \emph{xkcd.com}
\end{minipage}%
\\[2cm]
In computer vision research, the goal is to automatically extract information from an image.
In particular discerning semantic information, that is interpreting an image, is a prominent
research topic.
While much process has been made in recent years, computer vision systems still lack behind
humans in most tasks that require semantic information. Often these tasks can be formulated
in terms of \emph{semantic classes}, that is categories of parts, objects or scenes.
Examples include answering questions such as ``Is this a picture of a beach?'', ``How many
cars are in this image?'' or even ``What objects lie on the table?''.
These questions illustrate a range of possible tasks involving semantic categories,
from classifying images of single objects, over localization and counting of object classes,
to parsing a scene fully into objects and object classes, together with their relations.
%
While humans can distinguish tens of thousands of object classes, and have little trouble
in interpreting complex scenes, current methods are often restricted to a much smaller number
of classes, and only have limited capabilities to model interactions or relations.
We believe that context is one of the most important cues when it comes to classifying objects,
and therefore understanding scenes. Consequently, we target dense labeling of scenes on a fine-grained level.
The task of densely labeling image pixels into classes is called object class segmentation
or semantic segmentation.
\pagebreak

We choose this task in particular for several reasons:
\begin{itemize}
    \item Pixel-level annotation provides highly detailed information about the scene.
    \item Joint estimation of multiple classes allows for the use of \emph{context}.
    \item In contrast to category-agnostic segmentation approaches, object class segmentation
        has an unambiguous true labeling.
    \item A variety of manually annotated datasets is publicly available.
\end{itemize}
%
Applications of semantic segmentation and scene understanding include automatic
image interpretation for retrieval, autonomous driving and mobile robotics.
While these are already important applications, we expect that with the
abundance of camera data, and the proliferation of mobile computing, even
more applications will emerge soon.

We will in the following distinguish between the task of semantic segmentation,
which is usually distinguishing unstructured ``stuff'' classes such as road and
grass, and object class segmentation, which denotes the segmentation of very
structured classes, such as cars, planes and people.
We consider four different datasets in this thesis: the object class segmentation
datasets Graz02 and Pascal VOC 2010, and the semantic segmentation datasets MSRC-21,
and NYU V2.
%
Both tasks have the same ultimate goal of parsing, and therefore understanding, images
in terms of semantic classes, but usually employ  different mechanisms to represent and process.

Before we delve into semantic segmentation and object class recognition, we
will first investigate general grouping mechanisms in Chapter~\ref{ch:itm}.
Clustering is an important step in most semantic segmentation pipelines, in at
least two places: bottom-up segmentation and creation of dictionaries for
feature learning. Chapter~\ref{ch:itm} introduces a novel, information theoretic
algorithm that compares favourably with algorithms from the literature.
While we do not apply our algorithm to the task of bottom-up segmentation,
this is a promising avenue for future research.

One of the bottlenecks in learning object class segmentation and semantic
segmentation is the availability of training data.  While unlabeled image data,
and even data with semantic
``tags'' is available in practically unlimited quantities, semantic annotation
on pixel level is only available through laborious manual annotation. 
Chapter~\ref{ch:semi_supervised} introduces an approach to cope with this
shortcoming of object class segmentation approaches by introducing a method to
find segmentations automatically from image level annotations.

@@

\citet{nowozin2010parameter} provide a detailed evaluation of different aspects of learning object class
segmentation, that is somewhat orthogonal to this work. Their work considers the choice of features,
number superpixels and pairwise potentials for conditional maximum likelihood learning of tree-structured CRFs.
\citet{nowozin2010parameter} also compare conditional maximum likelihood learning with maximum margin learning,
finding little difference in accuracy.
We focus our work on the more popular neighborhood graphs, which do not allow for easy inference.
Therefore, conditional maximum likelihood learning would be much harder in our setting. For our studies,
we fix the considered feature set, and vary the maximum margin learning algorithm. Despite
this seemingly limited setup, we can make out large differences in the applicability of
some of the commonly used methods. For comparison, we also evaluate the learning algorithms
on chain CRFs, where inference is simple.
More recently, \citet{lucchi2013learning} proposed a novel algorithm for
efficiently learning structured prediction for semantic segmentation, using
approximate inference. In Chapter~\ref{ch:exact_learning}, we introduce an
algorithm that runs in similar time to the one proposed by
\citet{lucchi2013learning}, but is able to learn a CRF to the exact optimum on
the same dataset.

### 2013-09-19 main.tex
typo
use garamond, pick different sanse, leave math as is.
say everything twice. comit and delete lol

\usepackage{hyperref}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{calc}
\usepackage{refstyle}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{multicol}
%\usepackage{pslatex}
\usepackage{microtype}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{tabu}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{mdwlist}
\usepackage{url}
\usepackage[noend]{algpseudocode}
\usepackage[small]{caption}
\usepackage{setspace}
\usepackage{color}
\usepackage{minted}
\usepackage[german,english]{babel}

%\usepackage{tgpagella}
%\usepackage[T1]{fontenc}

%\usepackage{garamondx}
%\usepackage[urw-garamond]{mathdesign}
\usepackage{xspace}
\onehalfspacing
\KOMAoptions{DIV=last}

@@

The third and last part of this work investigates one of the most popular
approaches to the task of object class segmentation and semantic segmentation,
based on conditional random fields and structured prediction.
%
%For image labeling tasks, prediction using conditional random fields is usually
%only approximate, because of the presence of loops of the neighborhood graph in
%the the image. Consequently, learning in this setting is still little
%understood, despite the popularity of these methods.
%
We investigate several learning algorithms, in particular in combination with
approximate inference procedures. We show how structured models for image
segmentation can be learned exactly in practical settings, even in the presence
of many loops in the underlying neighborhood graphs.
The introduced methods provide results advancing the state-of-the art on two
complex benchmark datasets for semantic segmention, the MSRC-21 Dataset of RGB
images and the NYU V2 Dataset or RGB-D images of indoor scenes.

@@

During my work on this dissertation, my ideas were shaped by many of my fellow
students and researchers. First and foremost I am in debt to Hannes Schulz,
whose feedback proved to be invaluable during my studies.
I would like to thank Christoph Lampert for hosting me as a visiting researcher
at the IST Austria. His guidance and advice helped me in directing my further
research. My thanks are also extended to Sebastian Nowozin for his collaboration,
many helpful discussions, and advise.

@@

There are several directions for future research that we think would be
interesting to pursue as an extension of the presented results, in particular in
extending the second and third part:
\paragraph[Large-Scale Weakly Supervised Object Class Segmentation]
google
\paragraph[Inference Machines]
Recently \citet{stoyanov2011empirical} started a new trend in structured
prediction, which is sometimes called ``inference machines''. The basic
principle is simple: view the process of prediction with a given inference
procedure as a prediction method and optimize parameters of this method using
empirical risk minimization.  The work of \citet{stoyanov2011empirical}, used
loopy belief propagation as their inference algorithm and the optimization is
carried out simply using gradient descent on the non-convex but differentiable
loss function.  Other recent work in this direction includes
\citet{krahenbuhlparameter}, who used mean-field inference in a fully connected
conditional random field and \citet{jancsarylearning}, who used closed form
inference in a Gaussian CRF\@.
While these algorithms show great promise, their relation to the traditional
approach of structured prediction used in this work is mostly unclear. In particular,
if exact traditional learning is possible in a model, it is unclear how much
can be gained by direct empirical risk minimization.
Little empirical comparison was performed, and we are not aware of theoretical
work in this direction, leaving much room for future investigation of this
direction.
\paragraph[Non-Linear Models]
In this work, we only considered models that are linear in the input features--though
features are highly non-linear in the original input pixels.
Kernelization of structural support vector machines is trivial in theory, but
had only limited success in the context of CRFs for image
segmentation~\citep{lucchi2012structured}.
Two major alternatives for non-linear CRFs were proposed in the literature,
conditional neural fields~\citep{peng2009conditional} based on neural networks,
and decision tree fields~\citep{nowozin2011decision}, based on decision trees.
Conditional neural fields have only applied to sequence classification so far,
and extending them to our setting of semantic image segmentation would be very
interesting. Decision tree fields~(DTFs) on the other hand have been applied to loopy
graphs for image processing, but not for higher level tasks such as semantic
segmentation. If it is possible to include context in a meaningful way, it
might be possible to solve even object-centric tasks such as object class
segmentation with DTFs.

\paragraph[Latent Variable Models]
Something latent
\paragraph[Feature Design]
features are important
which low-level cues?
superpixels?

>>>>

\usepackage{hyperref}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{calc}
\usepackage{refstyle}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{multicol}
%\usepackage{pslatex}
\usepackage{microtype}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{tabu}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{mdwlist}
\usepackage{url}
\usepackage[noend]{algpseudocode}
\usepackage[small]{caption}
\usepackage{setspace}
\usepackage{color}
\usepackage{minted}
\usepackage[german,english]{babel}
%\renewcommand*\sfdefault{avant}
 %\usepackage{helvet}
\usepackage[defaultsans]{droidsans}
%\renewcommand*\sfdefault{lmssq}

\usepackage{garamondx}
%\usepackage[urw-garamond]{mathdesign}
\usepackage{xspace}
\onehalfspacing
\KOMAoptions{DIV=last}

@@

The third and main part of this work investigates one of the most popular
approaches to the task of object class segmentation and semantic segmentation,
based on conditional random fields and structured prediction.

We investigate several learning algorithms, in particular in combination with
approximate inference procedures. We show how structured models for image
segmentation can be learned exactly in practical settings, even in the presence
of many loops in the underlying neighborhood graphs.
The introduced methods provide results advancing the state-of-the art on two
complex benchmark datasets for semantic segmention, the MSRC-21 Dataset of RGB
images and the NYU V2 Dataset or RGB-D images of indoor scenes.

@@

During my work on this dissertation, my ideas were shaped by many of my fellow
students and researchers. First and foremost I am in debt to Hannes Schulz,
whose feedback proved to be invaluable during my studies.
I would like to thank Christoph Lampert for hosting me as a visiting researcher
at the IST Austria. His guidance and advice helped me in directing my further
research. My thanks are also extended to Sebastian Nowozin for his collaboration,
many helpful discussions, and advice.

@@

There are several directions for future research that we think would be
interesting to pursue as an extension of the presented results, in particular in
extending the second and third part:
\paragraph{Large-Scale Weakly Supervised Object Class Segmentation}
google
\paragraph{Inference Machines}
Recently \citet{stoyanov2011empirical} started a new trend in structured
prediction, which is sometimes called ``inference machines''. The basic
principle is simple: view the process of prediction with a given inference
procedure as a prediction method and optimize parameters of this method using
empirical risk minimization.  The work of \citet{stoyanov2011empirical}, used
loopy belief propagation as their inference algorithm and the optimization is
carried out simply using gradient descent on the non-convex but differentiable
loss function.  Other recent work in this direction includes
\citet{krahenbuhlparameter}, who used mean-field inference in a fully connected
conditional random field and \citet{jancsarylearning}, who used closed form
inference in a Gaussian CRF\@.
While these algorithms show great promise, their relation to the traditional
approach of structured prediction used in this work is mostly unclear. In particular,
if exact traditional learning is possible in a model, it is unclear how much
can be gained by direct empirical risk minimization.
Little empirical comparison was performed, and we are not aware of theoretical
work in this direction, leaving much room for future investigation of this
direction.
\paragraph{Non-Linear Models}
In this work, we only considered models that are linear in the input features--though
features are highly non-linear in the original input pixels.
Kernelization of structural support vector machines is trivial in theory, but
had only limited success in the context of CRFs for image
segmentation~\citep{lucchi2012structured}.
Two major alternatives for non-linear CRFs were proposed in the literature,
conditional neural fields~\citep{peng2009conditional} based on neural networks,
and decision tree fields~\citep{nowozin2011decision}, based on decision trees.
Conditional neural fields have only applied to sequence classification so far,
and extending them to our setting of semantic image segmentation would be very
interesting. Decision tree fields~(DTFs) on the other hand have been applied to loopy
graphs for image processing, but not for higher level tasks such as semantic
segmentation. If it is possible to include context in a meaningful way, it
might be possible to solve even object-centric tasks such as object class
segmentation with DTFs.

\paragraph{Latent Variable Models}
Something latent
\paragraph{Feature Design}
features are important
which low-level cues?
superpixels?

### 2013-09-19 nyu/paper.tex
say everything twice. comit and delete lol

\emph{Constant}
    A constant feature allows to model general neighborhood relations.

\emph{Color Contrast}
    We employ a non-linear color contrast, as is common in the computer
    vision literature, between the superpixel mean colors $c_i$ and $c_j$:
    \[
        \exp\left(-\gamma \lVert c_i - c_j \rVert^2\right).
     \]

\emph{Vertical Alignment}
    We model the directed angle between superpixel centers in the 2D image
    plane.  This allows the  model to learn that ``structure'' is above
    ``floor'', but not the other way around.

\emph{Depth Difference}
    We include the signed depth difference between superpixels, which
    allows the model to detect depth discontinuities that are not
    represented in the 2D neighborhood graph of the superpixels.

\emph{Normal Orientations}
    Differences in normal vector orientation are a strong clue on
    whether two superpixels belong to the same surface, and therefore the
    same structural class.
    We compute the 3D orientation of normals using the method of \citet{holz_2011_robocup},
    as implemented in the point cloud library (pcl)\footnote{\url{http://pointclouds.org}}.
    All normals within a superpixel are then averaged, to get a single orientation for each superpixel.
    The feature is computed as the difference of $\frac{\pi}{4}$ and the (undirected) angle between the normals belonging
    to two adjacent superpixels.  
    The feature is illustrated in \Figref{normal_feature}. The change
    in normal orientation highlights that pillow and wall are distinct
    objects, even though there is no strong distinction in color or depth.

>>>>

\paragraph{Constant}
    A constant feature allows to model general neighborhood relations.

\paragraph{Color Contrast}
    We employ a non-linear color contrast, as is common in the computer
    vision literature, between the superpixel mean colors $c_i$ and $c_j$:
    \[
        \exp\left(-\gamma \lVert c_i - c_j \rVert^2\right).
     \]

\paragraph{Vertical Alignment}
    We model the directed angle between superpixel centers in the 2D image
    plane.  This allows the  model to learn that ``structure'' is above
    ``floor'', but not the other way around.

\paragraph{Depth Difference}
    We include the signed depth difference between superpixels, which
    allows the model to detect depth discontinuities that are not
    represented in the 2D neighborhood graph of the superpixels.

\paragraph{Normal Orientations}
    Differences in normal vector orientation are a strong clue on
    whether two superpixels belong to the same surface, and therefore the
    same structural class.
    We compute the 3D orientation of normals using the method of \citet{holz_2011_robocup},
    as implemented in the point cloud library (pcl)\footnote{\url{http://pointclouds.org}}.
    All normals within a superpixel are then averaged, to get a single orientation for each superpixel.
    The feature is computed as the difference of $\frac{\pi}{4}$ and the (undirected) angle between the normals belonging
    to two adjacent superpixels.  
    The feature is illustrated in \Figref{normal_feature}. The change
    in normal orientation highlights that pillow and wall are distinct
    objects, even though there is no strong distinction in color or depth.
